<!-- question-type: prepare -->
### Exercise 3: The Price - Distance Relationship

By the end of this exercise, your plot should look similar to this one:

```{r}
#| fig-width: 5
#| fig-height: 5
#| fig-align: "center"
#| echo: false
ggplot(housing, 
       aes(x = distance, y = price)
       ) +
    geom_point(alpha = 0.05) +
    geom_smooth(method = "lm",
                se = FALSE, 
                linewidth = 1
                ) +
    labs(x = "Distance to CBD (km)", 
         y = "Price", 
         title = "Dwelling Price vs. Distance"
         )
```
**(a)** Why might we expect a relationship between price and distance to the CBD?

**(b)** Use the starter code below to build a scatter plot of **price vs. distance**, replacing `YOUR_VARIABLE_NAME` with the correct column names.

::: {.callout-tip title="What is a Scatter Plot?" collapse=true}
A **scatter plot** is used to visualize the relationship between two continuous variables by plotting individual data points. Each dot represents one observation, helping identify **patterns, clusters, and trends**. In this case, it allows us to explore how **house prices change with distance**.
:::

```{r}
#| eval: false
ggplot(housing, 
       aes(x = YOUR_VARIABLE_NAME, y = YOUR_VARIABLE_NAME)
       ) +
    geom_point()
```

**(c)** Since we have a large number of data points, many overlap, making it difficult to distinguish individual observations. Modify the `geom_point()` function to reduce over-plotting by adjusting the `alpha` value. Experiment with different values (e.g., 0.1, 0.05, 0.01) and describe how each affects the visualization.

```{r}
#| eval: false
ggplot(housing, 
       aes(x = YOUR_VARIABLE_NAME, y = YOUR_VARIABLE_NAME)
       ) +
    geom_point(alpha = YOUR_NUMBER)
```

**(d)** We can better understand the relationship by adding a statistical transformation to the plot. Add a fitted line using `geom_smooth(method = "lm", se = FALSE)`.

```{r}
#| eval: false
ggplot(housing, 
       aes(x = YOUR_VARIABLE_NAME, 
           y = YOUR_VARIABLE_NAME)
       ) +
    geom_point(alpha = YOUR_NUMBER) +
    YOUR_CODE_HERE
```

**(e)** Finally, add labels to the x-axis, y-axis, and title using the `labs()` function. Adjust the x- and y-axis scales as you deem appropriate.

```{r}
#| eval: false
ggplot(housing, 
       aes(x = YOUR_VARIABLE_NAME, y = YOUR_VARIABLE_NAME)
       ) +
    geom_point(alpha = YOUR_NUMBER) +
    YOUR_CODE_HERE +
    labs(x = "YOUR_LABEL", 
         y = "YOUR_LABEL", 
         title = "YOUR_TITLE"
         )
```

**(f)** What does the scatter plot reveal about the relationship between house prices and distance to the CBD? How does the line help interpret this relationship?


<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective** 
Students should:

- Create scatter plots to visualize the relationship between two continuous variables
- Use transparency (alpha) to manage overplotting in large datasets
- Add and interpret linear trend lines using `geom_smooth(method = "lm")`
- Format numeric axes appropriately for financial data
- Begin interpreting direction, strength, and limitations of observed relationships in data


✅   **Core Concepts to Highlight**

- Scatter plots are best for showing relationships between two numeric variables
- Overplotting can obscure patterns — using alpha makes dense regions clearer
- `geom_smooth(method = "lm")` fits a linear regression line; `se = FALSE` hides the shaded confidence interval
- The slope of the trend line shows direction of relationship; scatter around the line shows noise or variation
- The presence of outliers and spread highlights the limits of simple models and invites discussion of omitted variables

💬 **Suggested In-Class Prompts** (if needed)

“What are some reasons that a distant home might still have a high price?”

“What are we missing if we just look at this linear trend?”

“Could this relationship be non-linear? How would you test that?”

“What variables might improve our understanding if added to the model?”

📌 **Common Misunderstandings**

- Assuming the linear line implies causality — clarify it’s descriptive, not causal
- Not understanding what `geom_smooth(method = "lm")` does — thinking it's a smoothing curve by default
- Thinking the relationship is strong just because the trend line is visible — encourage students to consider variance and outliers


:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a). Why might we expect a relationship between price and distance to the CBD?**

In most cities, being closer to the Central Business District (CBD) is associated with higher property prices due to:

- Better access to jobs, transport, and amenities
- Scarcity of land and high demand in central locations
- Prestige or desirability of inner-city living

As distance from the CBD increases, we often expect prices to fall — although this relationship may not be perfectly linear due to other factors like transport links, school zones, or regional variation.

**(b)**

```{r}
ggplot(housing, 
       aes(x = distance, y = price)
       ) +
    geom_point()
```

**(c)**

```{r}
ggplot(housing, 
       aes(x = distance, y = price)
       ) +
    geom_point(alpha = 0.05)
```

Notes:

- alpha = 0.1: Points are semi-transparent — overlap is reduced but still busy.
- alpha = 0.05: Good balance — dense regions become visibly darker.
- alpha = 0.01: Almost all points are extremely faint; hard to see sparsely populated areas.

Lower alpha values reduce overplotting and help reveal density of data in crowded areas.

**(d)** 

```{r}
ggplot(housing, 
       aes(x = distance, y = price)
       ) +
    geom_point(alpha = 0.05) +
    geom_smooth(method = "lm", se = FALSE)
```

This adds a linear regression line, helping us see the overall trend in how price changes with distance.

**(e)**

```{r}
ggplot(housing, 
       aes(x = distance, y = price)
       ) +
    geom_point(alpha = 0.05) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
    labs(
        x = "Distance to CBD (km)", 
        y = "Price", 
        title = "House Price vs. Distance"
    )
```


**(f)**
The scatter plot shows a negative relationship between distance from the CBD and house price:

- As distance increases, price tends to decrease
- This is visible in the downward slope of the linear trend line
- There’s also substantial variation in prices at all distances — distance is a strong predictor, but not the only factor
- The line helps summarize the average trend, even in the presence of noisy data
- This supports the idea that proximity to the CBD carries a price premium, though other drivers of housing value are clearly at play

:::

:::

:::
<!-- END PROFILE:r-solutions -->