<!-- question-type: inclass -->
### Exercise 7: Putting the plots together

**(a).** Use the code below to combine the plots from the exercises above. Replace `PLOT_ONE`, `PLOT_TWO`, and `PLOT_THREE` with the plots you created in Exercises 4, 5, and 6.

```{r}
#| eval: false
PLOT_ONE /    
    (PLOT_TWO | PLOT_THREE) & 
    theme(legend.position = "none")
```

**(b).** Save this combined plot as a PDF file using the `ggsave()` function. You can specify the file name and dimensions using the `filename` and `width` and `height` arguments.

```{r}
#| eval: false
ggsave("YOUR_FILE_NAME", width = 12, height = 8)
```

<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective** 
Students should:

- Learn how to combine multiple ggplots into a single visual layout using `{patchwork}`
- Understand the basics of layout operators: `/` for rows, `|` for columns
- Practice naming plots as R objects, encouraging modular and readable code
- Use `ggsave()` to export a polished visualization for sharing or reporting

✅   **Core Concepts to Highlight**

- The {patchwork} package (built into recent versions of ggplot2) uses simple symbols:
    - `/` stacks plots vertically
    - `|` places plots side-by-side
- `theme(legend.position = "none")` removes an unnecessary legend
- `ggsave()` exports a plot to file — emphasize file name and dimensions are key
List them here


💬 **Suggested In-Class Prompts** (if needed)

“Why might you want to combine multiple plots into one?”

“What’s the difference between / and | in `{patchwork}`?”

“How would this be used in a business setting — e.g., for a report or dashboard?”

“What’s something you’d still want to tweak before sharing this with a wider audience?”
List as needed

📌 **Common Misunderstandings**

- Overwriting plots accidentally or misnaming them in the final composition
- Saving with a file extension other than .pdf without adjusting output type

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

```{r}
# Combine them into one layout
combined_plot <- price_hist / (region_bar | scttr) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Save the combined plot
ggsave("combined_price_plots.pdf", plot = combined_plot, width = 12, height = 8)
```


:::

:::

:::
<!-- END PROFILE:r-solutions -->