<!-- question-type: prepare -->
### Exercise 1: Understanding `ggplot2` Syntax

Consider the code below:

```{r}
#| eval: false

ggplot(housing, 
       aes(x = regionname)
       ) +
  geom_bar() +
  labs(
    x = "Region",
    y = "Count",
    title = "Number of Listings by Region"
  )
```

**(a)** What does `ggplot(housing, aes(x = regionname))` do?

**(b)** What is the role of `aes(x = regionname)` in this code?

**(c)** What does `geom_bar()` add to the plot?

**(d)** What does `labs()` do?

**(e)** In this code, what does the `+` operator do?

**(f)** Run the code above to produce the bar plot.

<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**
    
Students should be able to identify the main building blocks of a simple `ggplot2` command and explain how a plot is constructed step by step.

✅ **Core Concepts to Highlight**
    
* **`ggplot()` starts the plot**: It sets up the plotting object.
* **Data**: `housing` is the dataset being used.
* **Aesthetic mapping**: `aes(x = regionname)` maps a variable in the data to a visual feature of the plot.
* **Geometry**: `geom_bar()` tells `ggplot2` to draw bars.
* **Labels**: `labs()` adds descriptive text such as axis labels and a title.
* **Layering with `+`**: `ggplot2` plots are built by adding components together one layer at a time.

💬 **Suggested In-Class Prompts**
    
* What part of this code chooses the data?
* What part tells R what should go on the x-axis?
* Would this code draw anything if we removed `geom_bar()`?
* Why do we need the `+` operator in `ggplot2`?
* What changes if we edit `labs()`? What stays the same?
    
📌 **Common Misunderstandings**
    
* Students may think `ggplot()` by itself already draws the final graph. Emphasize that it sets up the plot, but the `geom_` layer is what actually draws something.
* Students may confuse `aes()` with labels. Clarify that `aes()` maps variables from the data, while `labs()` changes descriptive text shown on the figure.
* Students may not understand the `+` operator as part of a layered grammar. Stress that each `+` adds a new piece to the same plot.
* Some students may think `geom_bar()` requires them to manually count observations first. Clarify that `geom_bar()` automatically counts categories by default.


:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a)** `ggplot(housing, aes(x = regionname))` starts a plot using the housing dataset and maps the variable `regionname` to the x-axis.

**(b)** `aes(x = regionname)` defines the aesthetic mapping. It tells ggplot2 which variable from the dataset should be shown on the x-axis.

**(c)** `geom_bar()` adds bars to the plot. It creates a bar plot that counts how many observations fall into each category of regionname.

**(d)** `labs()` adds or changes labels on the plot, such as the x-axis label, y-axis label, and title.

**(e)** The `+` operator adds new layers or components to the plot. In ggplot2, we build plots step by step by adding pieces together.

**(f)** When you run the code, it produces a bar plot showing the number of listings in each region.

```{r}
#| eval: true
#| echo: false

ggplot(housing, 
       aes(x = regionname)
       ) +
  geom_bar() +
  labs(
    x = "Region",
    y = "Count",
    title = "Number of Listings by Region"
  )
```


:::

:::

:::
<!-- END PROFILE:r-solutions -->
