## The Business Challenge

```{r}
#| warning: false
#| echo: false
library(tidyverse)  # for plotting, includes ggplot
library(patchwork)  # for combining multiple plots into subfigures
library(scales)     # for formatting axis scales
library(ggokabeito) # color blind friendly color palette -- this course's default
```

```{r}
#| warning: false
#| eval: true
#| echo: false
#| cache: true

# Create data/ directory if it doesn't exist
if (!dir.exists("data")) {
  dir.create("data")
}
# Define URL and destination file path
url <- "https://github.com/unimelb-cmce-10002/fba-book/raw/refs/heads/main/data/melbourne_housing.csv"
destfile <- "data/melbourne_housing.csv"
# Download the file if it doesn't already exist
if (!file.exists(destfile)) {
  download.file(url, destfile, mode = "wb")
}
# Load the data
housing <- read_csv(destfile)
```

### The Topic: Understanding Melbourne’s Housing Market {.unnumbered}

Melbourne’s real estate market is one of the most dynamic in Australia. Property prices are influenced by several factors, including location, dwelling type, and proximity to the Central Business District (CBD). Buyers, sellers, and policymakers constantly analyze housing trends to make informed decisions. Understanding these trends is crucial in a $2 trillion+ housing market, where even small fluctuations in pricing expectations can have significant financial impacts.

A key question we will explore is: How do property prices vary by region, dwelling type, and location? By analyzing real estate data, we can identify patterns that help explain the drivers of Melbourne’s housing prices and their implications for different stakeholders in the market.

### The Data: The Melbourne Housing Market Dataset {.unnumbered}

We will use a dataset containing real estate sales data from Melbourne. The dataset provides insights into housing trends and includes key variables that influence property prices.

The dataset captures the property location within broader metropolitan regions (regionname), allowing us to compare trends across different areas. It also includes the type of dwelling, classified as House (h), Townhouse (t), or Apartment (u), enabling us to assess price differences based on housing type. The sale price (price) provides a measure of market value, while the distance to the CBD (distance) helps analyze the impact of proximity to the city center on property prices.

Similar datasets are widely used by real estate analysts, banks, and property investors to assess housing trends, predict future movements, and guide investment decisions. Through this tutorial, we’ll work with real-world data to uncover patterns and develop a deeper understanding of Melbourne’s property market. Let's dive in!

### Where we're headed {.unnumbered}

Just a few lines of R code transform numbers to data visualizations.

From this:

```{r}
#| echo: false 
housing %>%
    select(type, address, postcode, price, distance, regionname) %>%
    head(10)
```

to this:

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-align: "center"
#| echo: false

price_hist <- 
    ggplot(housing,
           aes(x = price, fill = type)
           ) +
    geom_histogram(
        bins = 50
    ) +
    # Log-scale the x-axis to handle price skew
    scale_x_continuous(
        trans = "log10",
        labels = scales::dollar
    ) +
    # Add labels and title
    labs(
        x = "Price", 
        y = "Number of Listings", 
        title = "Prices by Dwelling Type"
    ) +
    # Create separate plots for each dwelling type
    facet_wrap(~type) +
    # remove grey background
    theme_bw() +
    # Rotate x-axis labels for readability
    theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
    # Use colorblind-friendly fill palette
    scale_fill_okabe_ito() +
    # Center the title
    theme(
        plot.title = element_text(hjust = 0.5)
    )


scttr <-
    ggplot(housing,
           aes(x = distance, y = price)
           ) +
     # Scatter plot with reduced alpha to handle overplotting
    geom_point(alpha = 0.05, color = "grey") +
    # Polynomial fit to capture curvature (not just a straight line)
    geom_smooth(
        method = "lm",
        se = FALSE,
        color = "coral",
        linewidth = 1
    ) +
    # Format y-axis as dollars
    scale_y_continuous(labels = scales::dollar) +
    # Add labels and center the title
    labs(
        x = "Distance to CBD (km)",
        y = "Price",
        title = "Dwelling Price vs. Distance"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5)
    ) + 
    theme_bw() +
    theme(legend.position = "none")

region_bar <- 
    # 2. Create the plot
    ggplot(housing,
        # 2a. Reorder the region names based on frequency
        aes(x = fct_infreq(regionname))
    ) +
    # 3. Plot bar heights as proportions, not raw counts
    geom_bar(
      # THIS BIT IS HARD
      # YOU CAN NOT USE IT AND KEEP Y AXIS AS COUNTS
        aes(y = after_stat(count) / sum(after_stat(count))),
        fill = "grey"
    ) +
    # 4. Add informative axis labels and a title
    labs(
        x = "Region", 
        y = "Frequency",  # Note: this is technically a proportion, not % format
        title = "Number of Listings by Region"
    ) +
  # 5. Tweak the theme for better readability and presentation
    scale_y_continuous(labels = scales::percent)  + 
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 25, hjust = 1),  # Rotate x labels
        plot.title = element_text(hjust = 0.5)              # Center the title
    )


# Bringing together 
price_hist /    (region_bar | scttr)  & theme(legend.position = "none")


```

## Loading the Data

### R packages for today {.unnumbered}

```{r}
#| eval: false
library(tidyverse)  # for plotting, includes ggplot
library(patchwork)  # for combining multiple plots into subfigures
library(scales)     # for formatting axis scales
library(ggokabeito) # color blind friendly color palette -- this course's default
```

### Loading the Data in R {.unnumbered}

```{r}
#| warning: false
#| eval: false
#| echo: true
#| label: load-housing-data
#| cache: true

housing <-
    read_csv("data/melbourne_housing.csv")
```